---
title: "Association Rule Mining on DSS Demographics"
author: "_BI Innovators_ADSC41_"
date: "2025-04-22"
output: html_document
---


#Install packages
```{r}
library(arules)
library(arulesViz)
library(tidyverse)
```


#Read the data file  
```{r}
data_path <- "C:/Users/singhagiri/OneDrive/Desktop/assignment 2/dss-demographics-2021-sa2-december-2024.csv"
demo_data <- read.csv(data_path, header = TRUE, stringsAsFactors = TRUE)

```


#Inspect the dataset in R 
#Once the file has been imported to R, we often want to do few things to explore the dataset: 
```{r}
str(demo_data)
```

```{r}
summary(demo_data)
```

```{r}
names(demo_data)
head(demo_data)
tail(demo_data)
```


#Check the dimension 
```{r}
dim(demo_data)
```

#check column names
```{r}
colnames(demo_data)

```


#Plot and explore the  “demo_data” dataset  with barplot() function  
```{r}
# Load required packages
library(tidyverse)

# Load the dataset
data <- read.csv("C:/Users/singhagiri/OneDrive/Desktop/assignment 2/dss-demographics-2021-sa2-december-2024.csv", stringsAsFactors = FALSE)

# View column names (for reference)
colnames(data)

# Step 1: Select only benefit columns (columns 4 to 24 based on your list)
benefits_data <- data[, 4:24]

# Step 2: Convert numeric values to "Yes"/"No" (Yes if benefit count > 0)
marketbasket <- as.data.frame(lapply(benefits_data, function(x) ifelse(x > 0, "Yes", "No")))

# Step 3: Use colSums to count Yes and No for each benefit
yes <- colSums(marketbasket == "Yes")
no  <- colSums(marketbasket == "No")

# Step 4: Combine into a matrix for plotting
purchased <- rbind(Yes = as.numeric(yes), No = as.numeric(no))
colnames(purchased) <- names(yes)

# Step 5: Plot 1 - Stacked Barplot
barplot(purchased,
        legend = rownames(purchased),
        main = "Distribution of Benefits Received (Stacked)",
        col = c("skyblue", "salmon"),
        las = 2,              # Rotate x labels
        cex.names = 0.7)      # Adjust label size

# Step 6: Plot 2 - Side-by-Side Barplot
barplot(purchased,
        beside = TRUE,
        legend = rownames(purchased),
        main = "Distribution of Benefits Received (Side-by-Side)",
        col = c("skyblue", "salmon"),
        las = 2,
        cex.names = 0.7)

```

#create Association rules .  
#Usage of apriori () function 
```{r}
rules <- apriori(demo_data)
```


#Get the summary of these rules  
```{r}
summary(rules) 
```



#To inspect the rules please use the following line of code:  
#Since there are too many rules we need to reduce them into smaller number of rules hence 
we have to specify the parameters using the following code:  
#Complete R Code Using Your Dataset and Specified Apriori Parameters

```{r}
#Apply Apriori algorithm with your specified parameters
rules <- apriori(demo_data,
                 parameter = list(minlen = 2,
                                  maxlen = 3,
                                  conf = 0.95))

# Step 6: View summary and top 10 rules
summary(rules)
inspect(sort(rules, by = "confidence")[1:10])
```


#To Visualize these rules we can use a package called “arulesViz”  
# Plot the rules. 

```{r}
install.packages(arulesViz)  
library(arulesViz)   
plot(rules)  
```

#Plot the rules in groups:
```{r}
plot(rules, method="grouped") 
```

#Below code displays a scatterplot matrix to compare the support, confidence, and lift 
```{r}
plot(rules@quality)  
```



```{r}
# Visualize rules as a graph
plot(rules, method = "graph", engine = "htmlwidget")

# Optional: matrix or grouped visual
plot(rules, method = "grouped")

```


#ruleExplorer() function. 
```{r}
rules_ex <-apriori(demo_data,  
parameter =list(minlen=2,maxlen=4,conf=0.75))
```

```{r}
ruleExplorer(rules_ex)  
```

```{r}
ruleExplorer(demo_data)  
```




