---
title: "Logistic Regression on DSS Demographics Data"
author: "_BI Innovators_ADSC41_"
date: "2025-04-22"
output: html_document
---
```{r}
library(ggplot2)
library(dplyr)
library(corrplot)
library(GGally)
library(readr)
```

```{r}
# Load the dataset 
house <- read_csv("dss-demographics-2021-sa2-december-2024.csv")
```
##.Simple and Multiple Linear Regression Models

```{r}
# Preview the data
head(house)

```

```{r}
dim(house)

```
```{r}
str(house)

```


```{r}
# Linear regression
house.lm1 <- lm(`JobSeeker Payment` ~ `Age Pension`, data = house)
summary(house.lm1)

```

```{r}
ggplot(house, aes(x = `Age Pension`, y = `JobSeeker Payment`)) +
  geom_point(color = 'blue', size = 6) +
  geom_smooth(method = lm, se = FALSE, color = 'red') +
  labs(
    title = "JobSeeker Payment vs Age Pension",
    x = "Age Pension",
    y = "JobSeeker Payment"
  ) +
  theme_minimal()

```

```{r}
# Add residuals to the original dataset
house$residuals <- resid(house.lm1)

# Plot residuals vs fitted values
ggplot(house, aes(x = `Age Pension`, y = residuals)) +
  geom_point(color = 'darkgreen', size = 4, alpha = 0.7) +
  geom_hline(yintercept = 0, color = 'red', linetype = "dashed") +
  labs(
    title = "Residual Plot: JobSeeker Payment ~ Age Pension",
    x = "Age Pension",
    y = "Residuals"
  ) +
  theme_minimal()

```
```{r}
# Multiple Linear Regression
multi.lm <- lm(`JobSeeker Payment` ~ `Age Pension` + `Disability Support Pension` + `Carer Payment` + `Family Tax Benefit A`, data = house)

# Summary of the model
summary(multi.lm)

```
```{r}
# Add fitted values and residuals
house$fitted <- fitted(multi.lm)
house$residuals <- resid(multi.lm)

# Residual plot
ggplot(house, aes(x = fitted, y = residuals)) +
  geom_point(color = "purple", size = 4, alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Residuals vs Fitted Values (Multiple Linear Regression)",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

```
##.Logistic regression models

```{r}
# Remove all rows with missing (NA) values
house_clean <- na.omit(as.data.frame(house))
```

```{r}
# Show summary statistics
summary(house_clean)
```

```{r}
# Dimensions: number of rows and columns
dim(house_clean)

# Structure of the dataset (column types, etc.)
str(house_clean)

# Preview first few rows
head(house_clean)
```

```{r}
colnames(house)
```
##.Correlogram
```{r}
library(corrplot)
```

```{r}
# Read the dataset
house <- read.csv("dss-demographics-2021-sa2-december-2024.csv")

```
```{r}
# Subset only numeric-relevant columns (from column 4 to 24)
house_numeric <- house[, 4:24]

# Convert all columns to numeric (force conversion)
house_numeric <- as.data.frame(lapply(house_numeric, function(x) as.numeric(as.character(x))))

# Remove rows with NA (optional but recommended)
house_numeric <- na.omit(house_numeric)

# Compute correlation matrix
cor_matrix <- cor(house_numeric)
```

```{r}
# Visualize with corrplot
install.packages("corrplot")
library(corrplot)

corrplot(cor_matrix, method = "color",
         type = "upper",
         addCoef.col = "black",
         tl.col = "black", tl.srt = 45,
         col = colorRampPalette(c("red", "white", "blue"))(200),
         number.cex = 0.7)
```
##.Plot 1: ggpairs() — Pairwise Scatterplots for All Variables
```{r}
ggpairs(data = house_numeric, title = "Pairwise Comparison of Social Support Variables")

```


##.Linear Regression Model 1: JobSeeker vs Age Pension
```{r}
model1 <- lm(JobSeeker.Payment ~ Age.Pension, data = house_numeric)
summary(model1)

# Plot
ggplot(house_numeric, aes(x = Age.Pension, y = JobSeeker.Payment)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("JobSeeker Payment vs Age Pension")

```
##. Linear Regression Model 2: Disability Support vs Carer Payment
```{r}
model2 <- lm(Disability.Support.Pension ~ Carer.Payment, data = house_numeric)
summary(model2)

# Plot
ggplot(house_numeric, aes(x = Carer.Payment, y = Disability.Support.Pension)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle("Disability Support vs Carer Payment")

```
##.Logistic regression models

```{r}


# Convert relevant columns to numeric
house_numeric <- house[, 4:24]
house_numeric <- as.data.frame(lapply(house_numeric, function(x) as.numeric(as.character(x))))

# Identify complete cases and filter both datasets
complete_rows <- complete.cases(house_numeric)
house_numeric <- house_numeric[complete_rows, ]
house_filtered <- house[complete_rows, ]

# Combine safely
house_full <- cbind(house_filtered[, 1:3], house_numeric)

```
##.Create a Binary Dependent Variable
```{r}
# Create binary outcome variable
median_jobseeker <- median(house_numeric$JobSeeker.Payment, na.rm = TRUE)
house_full$high_jobseeker <- ifelse(house_full$JobSeeker.Payment > median_jobseeker, 1, 0)

# Convert to factor
house_full$high_jobseeker <- as.factor(house_full$high_jobseeker)

```

##.Split into Training & Validation Sets
```{r}
set.seed(123)
pd <- sample(2, nrow(house_full), replace = TRUE, prob = c(0.7, 0.3))
train <- house_full[pd == 1, ]
validate <- house_full[pd == 2, ]

```

##.Build Logistic Regression Model
```{r}
model_all <- glm(high_jobseeker ~ ., data = train[, -c(1,2,3)], family = "binomial")  # Exclude date, code, and name
summary(model_all)

```
##.Model Evaluation – Training Set
```{r}
# Predict on training set
train_pred <- ifelse(predict(model_all, type = "response") > 0.5, 1, 0)
train_tab <- table(predicted = train_pred, actual = train$high_jobseeker)
train_tab

# Accuracy
sum(diag(train_tab)) / sum(train_tab)

```
##.Model Evaluation – Validation Set
```{r}
# Predict on validation set
validate_pred <- ifelse(predict(model_all, newdata = validate, type = "response") > 0.5, 1, 0)
validate_tab <- table(predicted = validate_pred, actual = validate$high_jobseeker)
validate_tab

# Accuracy
sum(diag(validate_tab)) / sum(validate_tab)

```
##.Plot Logistic Regression Fit
```{r}
library(ggplot2)
ggplot(train, aes(x = Age.Pension, y = as.numeric(as.character(high_jobseeker)))) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), color = "blue") +
  labs(title = "Logistic Regression: High JobSeeker ~ Age Pension", y = "Probability of High JobSeeker")

```


